---
title: "Week 38 CRAN packages"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Using Quarto to do data exploration

Since I have not really played with .qmd file... I'm just trying to utilize instead of just scribbling down on r scripting file.

## Installing Libraries

I didn't realize there's tidytuesdayR package to get the data swiftly from official repo on Github till today!

```{r}

library(tidyverse)
library(janitor)
library(ggforce)
library(tidygraph)
library(ggraph)
library(tidytext)
library(lubridate)
library(gt)
library(gtExtras)
```

## Getting the Data Set for Week 38 2023

Loading up the dataset and augmenting few things

```{r}

tuesdata <- tidytuesdayR::tt_load('2023-09-19')
names(tuesdata)
## OR
#tuesdata <- tidytuesdayR::tt_load(2023, week = 38)

cran <- tuesdata$cran_20230905 |>
  clean_names() 

cran <- cran |> 
  mutate(url = str_glue("https://CRAN.R-project.org/package={package}"),
         package_url = str_glue("<a href='{url}'> {package} </a>"),
         package_initial=str_to_upper(str_sub(package,1L,1L)),
         package_name_len=str_count(package))

authors <- tuesdata$package_authors |> 
  clean_names()

nodes <- tuesdata$cran_graph_nodes |>
  clean_names()

edges <- tuesdata$cran_graph_edges |>
  clean_names()

# Reconstruct a graph
cran_graph <- tbl_graph(
  nodes = nodes, edges = edges, directed = FALSE
)

cran_graph <- cran_graph |>
  mutate(idx=row_number())

cran_graph |> filter(str_detect(name,"Wickham"))

cran_graph |> as_tibble() |>
  count(dist2hw)


cran |> skimr::skim()

### Package with Long Names! 
cran |> arrange(desc(package_name_len))  |>
  select(package_url, description, title, published) |>
  head(30) |>
  gt() |>
  fmt_markdown(package_url)
```

## Words used in Title & Description

Just trying to get sense of data set. I don't know what I want to visualize yet :D

```{r}

names(cran)

cran_detail <-cran |>
  arrange(desc(published)) |>
  group_by(package_url,package, package_initial) |>
  summarise(title=first(title),
            description=first(description),
            last_published = max(published, na.rm=T))  |>
  ungroup() |>
  mutate(sentences = str_c(title, " ", description))


cran_words <- cran_detail |>
  unnest_tokens(output="word",input="sentences") |>
  anti_join(stop_words) 

cran_words_by_auth <-cran_words |> count(word,package,sort=T) |>
  inner_join(authors |> group_by(package) |>
               summarise(authors_list=list(authors_r),
                         authors_cnt=n())) |>
  unnest(authors_list) |>
  mutate(n=n/authors_cnt) |> ## if package is co-authoured then we should weigh them?
  count(authors_list,word,wt=n,sort=T) |>
  bind_tf_idf(term=word,document=authors_list,n=n)


cran_words_by_auth |> filter(str_detect(word,"tuesday")) |>
  arrange(desc(tf_idf))

cran |> filter(str_detect(str_c(title," ", description),
                           regex("tidytuesday", ignore_case=T))) |>
  select(package_url,title,description,published,authors_r) |>
  gt() |>
  fmt_markdown(package_url)
```

## CRAN package dealing with Japan

Surprise to see 31 packages that includes word "japan" in description. Hmm, but knowing there's close to 20K packages out there. 31 is maybe too little...

```{r}
cran_japan <- cran |>
  filter(str_detect(paste(title,description,package),regex("(japan|nihon)", ignore_case=T)))

cran_japan |>
  select(package_url,title,description,author, published) |>
  gt() |>
  fmt_markdown(package_url)
```

## CRAN package dealing with Color Palette in some way :)

Because I am typically obsessed with color palette in general. I maybe able to discover some new packages, that I haven't heard of yet! I'm looking for any package with either title or description including "colour" and "palette" to narrow down. There are 104 packages. I think it's good amount to do some network visualization. :)

```{r}
cran_color <- cran |>
  filter(str_detect(paste(title,description,package),regex("(colour|color)", ignore_case=T)) & str_detect(paste(title,description,package),regex("(palette)", ignore_case=T)))

cran_color |>
  select(package_url,title,description,author, published) |>
  gt() |>
  fmt_markdown(package_url)


```

## CRAN Packages Dependency

I'd say package has dependency if package name is listed under depends column or imports column. I'm not sure the difference between the 2. Depends sounds like it should only be listing R version or package that are pre-installed with R, where as imports would be other packages on CRAN?!

I've googled and I have gotten following explanation, but now I'm confused to what it means when package is "attached" . 🤔\
\
`"The main difference is that a package you list in Imports will just be loaded when you use it, whereas a package you list in Depends will be attached when your package is attached. Unless there is a good reason otherwise, you should always list packages in Imports not Depends ."  (R Packages (2e) Chapter 10 Dependencies: Mindset and Background)`

```{r}

### grrr.. yes I should try to tidy below up bit....  
cran_dependency <- cran |> select(package,depends,imports) |>
  mutate(depends=replace_na(depends,""),
         imports=replace_na(imports,"")) |>
  mutate(relates=str_trim(str_c(depends,", ",imports))) |>
  mutate(relates=str_split(relates,",")) |>
  unnest(relates) |>
  mutate(relates=str_trim(relates)) |>
  filter(str_count(relates)>1) |>
  filter(!str_detect(relates,"R (.+)")) |>
  mutate(relates=str_trim(str_remove_all(relates,"\\(.+\\)"))) |>
  ### I just realized that when you use . it means any character except \n
  mutate(relates=str_trim(str_remove_all(relates,"\\((\\\n|.)+\\)"))) |> 
  filter(!str_detect(relates,"^R$"))

### Create edge lists  
cran_color_e <-cran_dependency |> 
  filter(package %in% cran_color$package) |>
  select(package,relates) |> distinct()

###. Create node lists
cran_color_n <- tibble(
  name=c(cran_color_e$package,cran_color_e$relates) |> unique()
  ) |> left_join(cran_detail |> select(name=package,title,description,last_published)) |>
  mutate(in_package=name %in% cran_color_e$package)

### relates_pkg is a flag to say if dependent package also appears as packages of interest
cran_color_e <-cran_color_e |> mutate(relates_pkg = relates %in% cran_color_e$package)

### Use the edges and nodes and make them into tbl_graph object
cran_color_g <- tbl_graph(nodes=cran_color_n,edges=cran_color_e, directed=F)


### Just experimenting with group_xxx. 
#### group_infomap = Infomap community finding = find community structure that minimizes the expected description length of a random walker trajectory...  hmmm.  
#### group_edge_betweenness = Many networks consist of modules which are densely connected themselves but sparsely connected to other modules.  hmmm..  

cran_color_g <- cran_color_g |>
  mutate(grp_infomap=group_infomap(),
         grp_edge_btwn=group_edge_betweenness(),
         deg=centrality_degree())

### with infomap how did package clustered into group?
cran_color_g |> as_tibble() |>
  group_by(grp_infomap) |>
  summarise(pkgs=paste(name,collapse=", "),
            cnt=n()) 

### Original Layout 
#### I don't know the best way to pick layout type, I've tried kk, fr, stress, nicely, lgl etc.  I like Sugiyama the best for this.  
cran_color_g |>
  ggraph("sugiyama") +
  geom_edge_diagonal(edge_width=0.1) +
  geom_node_point(aes(color=in_package)) +
  geom_node_text(aes(label=name), family="Roboto Condensed", hjust="outwards") +
  coord_flip()

### I want to save the coordinate because I want to use the layout as a base. 
cran_color_layout <- cran_color_g |>
  create_layout("sugiyama")

###  I'm brute-forcefully adjusting layout...  

cran_color_layout2 <-cran_color_layout |>
  group_by(y) |>
  mutate(x = scales::rescale(row_number(x), to=range(x))) |>
  mutate(y=if_else(name=="ggplot2",5.8,y)) |>
  mutate(x=if_else(name=="earthtones",107,x)) |>
  mutate(x=if_else(name %in% c("earthtones","echogram","ggmap","geosphere","readHAC"),x-112,x)) |>
  mutate(x=if_else(name=="ggplot2",100,x)) |>
  ungroup()

cran_color_e |> filter(relates=="ggplot2") |> pull(package) -> ggplot2_grp
cran_color_e |> 
  filter(!package %in% (cran_color_e |> filter(relates=="ggplot2") |> pull(package))) |>
  pull(package)|> unique() -> not_ggplot2_grp

#pal_mat <- ggthemes::tableau_color_pal("Hue Circle")(19) |> sample(size=19) |> farver::decode_colour() 

# c("007f5f-2b9348-55a630-80b918-aacc00-bfd200-d4d700-dddf00-eeef20-ffff3f") |>
#   str_split("-") |>
#   unlist() -> col_chi

ggthemes::tableau_color_pal("Tableau 10")(10) -> col_chi

pal_mat <- col_chi |> farver::decode_colour()
col_chi|> scales::show_col()
  
cran_color_layout$x |> range()

cran_color_g |>
  mutate(deg=centrality_degree()) |>
  as_tibble() |> 
  count(deg) |>
  pull(deg) -> possible_deg_cnt


draw_by_grp <- function(deg_cnt,...){
  cran_color_g |>
  activate(nodes) |>
  mutate(x=cran_color_layout2$x, # use cran_color_layout if we want original
         y=cran_color_layout2$y,
         deg = centrality_degree(),
         color=colourvalues::color_values(deg,pal_mat)
         ) |>
  arrange(in_package,desc(deg)) |>
  #filter(in_package)|>
  #filter(!in_package) |>
  filter(deg > deg_cnt)|>
  ggraph(layout="manual",x=x,y=y) +
  geom_edge_diagonal(edge_width=0.3, 
                     aes(color=I(.N()$color[to]),
                         start_cap = label_rect(node1.name,fontface="bold",family="Roboto Condensed"), 
                         end_cap = label_rect(node2.name, fontface="bold", family="Roboto Condensed")))+
  geom_node_label(aes(fill=I(if_else(in_package,color,colorspace::darken(color,amount=0.85))),
                      label=name), 
                  color="transparent") +
  geom_node_text(aes(label=name, fontface=if_else(in_package,"bold","italic"),
                     color=I(if_else(in_package,"#000000de","#ffffffde"))), 
                 family="Roboto Condensed") +
  coord_flip(ylim=c(0,6),xlim=c(-10,112)) 
  
  #ggsave(str_glue("2023/week-38/output/cran_color_sugiyama_{deg_cnt}.png"), device=ragg::agg_png, width=9*1.2, height=12*1.2)
  
  ggsave(str_glue("2023/week-38/output/cran_color_sugiyama_{deg_cnt}.svg"), device=svglite::svglite, width=9*1.2, height=12*1.2)
  
  
}


draw_by_grp(deg_cnt=0)

### draw by degree and produce separate images so I can combine them to make animation
possible_deg_cnt <- unname(possible_deg_cnt)
possible_deg_cnt[c(1:20)] |> rev() |> map(draw_by_grp)

### I feel like there's just so much efficient way to make animation.. but I can't figure out how just yet...


```

### Creating Table listing out 104 CRAN packages dealing with Colour Palette

Trying to figure out how to use package gt :) https://gt.rstudio.com/reference/index.html

```{r}

cran_color |> 
  left_join(cran_color_g |> as_tibble() |> select(package=name,deg) |>
              mutate(deg_color = colourvalues::color_values(deg,pal_mat))) |>
  select(deg,deg_color,package_url,package,published,title,description,depends,imports) |>
  mutate(package=str_glue("<h4><span style='background-color:{deg_color};color:{deg_color}'> ■ </span>  {package_url} 📦 </h4>")) |>
  arrange(deg) |>
  mutate(deg=if_else(is.na(deg),"Package with No Depenency - Not Included in Network Visualization",if_else(deg==1,"Package with 1 Dependency",str_c("Package with ",deg, " Dependencies")))) |>
  group_by(deg) |>
  gt() |>
  fmt_markdown(c(package)) |> 
  cols_hide(c(package_url, deg_color)) |>
  gt_theme_538() |> 
  tab_header(title="CRAN Package dealing with Colour Palettes",
             subtitle="Week 38 TidyTuesday Data Visualization Challenge") |>
  gtsave("2023/week-38/output/cran_colour_palette_packages.html")


```

### Creating Table Listing Package Names in Visuals for README

If you create README.md under sub-folders within your main Repository, it does get rendered nicely as README file on main repository page. Which I only realized now.

I wanted to list out name of Packages that were shown in visuals in table format in README.md

I've generated table with below code and copied over to README.md for now.

```{r}

cran_color |> 
  left_join(cran_color_g |> as_tibble() |> select(package=name,deg) |>
              mutate(deg_color = colourvalues::color_values(deg,pal_mat))) |>
  select(deg,deg_color,package_url,package,published,title,description,depends,imports) |>
  mutate(package=str_glue("{package_url} 📦")) |>
  arrange(deg) |>
  mutate(deg=if_else(is.na(deg),"Package with No Depenency - Not Included in Network Visualization",if_else(deg==1,"Package with 1 Dependency",str_c("Package with ",deg, " Dependencies")))) |>
  group_by(deg) |>
  summarise(package_names=paste(package, collapse=" / ")) %>%
  knitr::kable("markdown") %>% clipr::write_clip()
```

### Wordcloud2 of Colour Palette Packages on CRAN

```{r}

cran_words |> filter(package %in% cran_color$package) |>
  count(word,sort=T) |>
  head(100) |>
  mutate(n=sqrt(n)) |>
  wordcloud2::wordcloud2(rotateRatio=0, fontFamily="Roboto Condensed",
                         color=sample(col_chi,size=100,replace=T))
```
