---
title: "Week 38 CRAN packages"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Using Quarto to do data exploration

## Installing Libraries

I didn't realize there's tidytuesdayR package to get the data swiftly from official repo on Github till today.

```{r}

library(tidyverse)
library(janitor)
library(ggforce)
library(tidygraph)
library(ggraph)
library(tidytext)
library(lubridate)
library(gt)
library(gtExtras)
```

## Getting the Data Set for Week 38 2023

Loading up the dataset and augmenting few things

```{r}

tuesdata <- tidytuesdayR::tt_load('2023-09-19')
names(tuesdata)
## OR
#tuesdata <- tidytuesdayR::tt_load(2023, week = 38)

cran <- tuesdata$cran_20230905 |>
  clean_names() 

cran <- cran |> 
  mutate(url = str_glue("https://CRAN.R-project.org/package={package}"),
         package_url = str_glue("<a href='{url}'> {package} </a>"),
         package_initial=str_to_upper(str_sub(package,1L,1L)),
         package_name_len=str_count(package))

authors <- tuesdata$package_authors |> 
  clean_names()

nodes <- tuesdata$cran_graph_nodes |>
  clean_names()

edges <- tuesdata$cran_graph_edges |>
  clean_names()

# Reconstruct a graph
cran_graph <- tbl_graph(
  nodes = nodes, edges = edges, directed = FALSE
)

cran_graph <- cran_graph |>
  mutate(idx=row_number())

cran_graph |> filter(str_detect(name,"Wickham"))

cran_graph |> as_tibble() |>
  count(dist2hw)


cran |> skimr::skim()

### Package with Long Names! 
cran |> arrange(desc(package_name_len))  |>
  select(package_url, description, title, published) |>
  head(30) |>
  gt() |>
  fmt_markdown(package_url)
```

## Words used in Title & Description 

```{r}

names(cran)

cran_detail <-cran |>
  arrange(desc(published)) |>
  group_by(package_url,package, package_initial) |>
  summarise(title=first(title),
            description=first(description),
            last_published = max(published, na.rm=T))  |>
  ungroup() |>
  mutate(sentences = str_c(title, " ", description))


cran_words <- cran_detail |>
  unnest_tokens(output="word",input="sentences") |>
  anti_join(stop_words) 

cran_words_by_auth <-cran_words |> count(word,package,sort=T) |>
  inner_join(authors |> group_by(package) |>
               summarise(authors_list=list(authors_r),
                         authors_cnt=n())) |>
  unnest(authors_list) |>
  mutate(n=n/authors_cnt) |> ## if package is co-authoured then we should weigh them?
  count(authors_list,word,wt=n,sort=T) |>
  bind_tf_idf(term=word,document=authors_list,n=n)


cran_words_by_auth |> filter(str_detect(word,"tuesday")) |>
  arrange(desc(tf_idf))

cran |> filter(str_detect(str_c(title," ", description),
                           regex("tidytuesday", ignore_case=T))) |>
  select(package_url,title,description,published,authors_r) |>
  gt() |>
  fmt_markdown(package_url)
```

## CRAN package dealing with Japan 

```{r}
cran_japan <- cran |>
  filter(str_detect(paste(title,description,package),regex("(japan|nihon)", ignore_case=T)))

cran_japan |>
  select(package_url,title,description,author, published) |>
  gt() |>
  fmt_markdown(package_url)
```

## CRAN package dealing with Color Palette in some way :)  

```{r}
cran_color <- cran |>
  filter(str_detect(paste(title,description,package),regex("(colour|color)", ignore_case=T)) & str_detect(paste(title,description,package),regex("(palette)", ignore_case=T)))

cran_color |>
  select(package_url,title,description,author, published) |>
  gt() |>
  fmt_markdown(package_url)


```

## CRAN Packages Dependency

```{r}
cran_dependency <- cran |> select(package,depends,imports) |>
  mutate(depends=replace_na(depends,""),
         imports=replace_na(imports,"")) |>
  mutate(relates=str_trim(str_c(depends,", ",imports))) |>
  mutate(relates=str_split(relates,",")) |>
  unnest(relates) |>
  mutate(relates=str_trim(relates)) |>
  filter(str_count(relates)>1) |>
  filter(!str_detect(relates,"R (.+)")) |>
  mutate(relates=str_trim(str_remove_all(relates,"\\(.+\\)"))) |>
  mutate(relates=str_trim(str_remove_all(relates,"\\((\\\n|.)+\\)"))) |>
  filter(!str_detect(relates,"^R$"))


cran_color_e <-cran_dependency |> 
  filter(package %in% cran_color$package) |>
  select(package,relates) |> distinct()

cran_color_n <- tibble(
  name=c(cran_color_e$package,cran_color_e$relates) |> unique()
  ) |> left_join(cran_detail |> select(name=package,title,description,last_published)) |>
  mutate(in_package=name %in% cran_color_e$package)

cran_color_e <-cran_color_e |> mutate(relates_pkg = relates %in% cran_color_e$package)

cran_color_g <- tbl_graph(nodes=cran_color_n,edges=cran_color_e, directed=F)

cran_color_g <- cran_color_g |>
  mutate(grp_infomap=group_infomap(),
         grp_edge_btwn=group_edge_betweenness(),
         deg=centrality_degree())

cran_color_g |> as_tibble() |>
  group_by(grp_infomap) |>
  summarise(pkgs=paste(name,collapse=", "),
            cnt=n()) 

### Original Layout 
cran_color_g |>
  ggraph("sugiyama") +
  geom_edge_diagonal(edge_width=0.1) +
  geom_node_point(aes(color=in_package)) +
  geom_node_text(aes(label=name), family="Roboto Condensed", hjust="outwards") +
  coord_flip()

cran_color_layout <- cran_color_g |>
  create_layout("sugiyama")



### It's too crowded.  I'm brute-forcefully adjusting layout...
cran_color_layout2 <-cran_color_layout |>
  group_by(y) |>
  mutate(x = scales::rescale(row_number(x), to=range(x))) |>
  mutate(y=if_else(name=="ggplot2",5.8,y)) |>
  mutate(x=if_else(name=="earthtones",107,x)) |>
  mutate(x=if_else(name %in% c("earthtones","echogram","ggmap","geosphere","readHAC"),x-112,x)) |>
  mutate(x=if_else(name=="ggplot2",100,x)) |>
  ungroup()

cran_color_e |> filter(relates=="ggplot2") |> pull(package) -> ggplot2_grp
cran_color_e |> 
  filter(!package %in% (cran_color_e |> filter(relates=="ggplot2") |> pull(package))) |>
  pull(package)|> unique() -> not_ggplot2_grp

#pal_mat <- ggthemes::tableau_color_pal("Hue Circle")(19) |> sample(size=19) |> farver::decode_colour() 

# c("007f5f-2b9348-55a630-80b918-aacc00-bfd200-d4d700-dddf00-eeef20-ffff3f") |>
#   str_split("-") |>
#   unlist() -> col_chi

ggthemes::tableau_color_pal("Tableau 10")(10) -> col_chi

pal_mat <- col_chi |> farver::decode_colour()
col_chi|> scales::show_col()
  

cran_color_layout$x |> range()



cran_color_g |>
  mutate(deg=centrality_degree()) |>
  as_tibble() |> 
  count(deg) |>
  pull(deg) -> possible_deg_cnt


draw_by_grp <- function(deg_cnt,...){
  cran_color_g |>
  activate(nodes) |>
  mutate(x=cran_color_layout2$x, # use cran_color_layout if we want original
         y=cran_color_layout2$y,
         deg = centrality_degree(),
         color=colourvalues::color_values(deg,pal_mat)
         ) |>
  arrange(in_package,desc(deg)) |>
  #filter(in_package)|>
  #filter(!in_package) |>
  filter(deg > deg_cnt)|>
  ggraph(layout="manual",x=x,y=y) +
  geom_edge_diagonal(edge_width=0.3, 
                     aes(color=I(.N()$color[to]),
                         start_cap = label_rect(node1.name,fontface="bold",family="Roboto Condensed"), 
                         end_cap = label_rect(node2.name, fontface="bold", family="Roboto Condensed")))+
  geom_node_label(aes(fill=I(if_else(in_package,color,colorspace::darken(color,amount=0.85))),
                      label=name), 
                  color="transparent") +
  geom_node_text(aes(label=name, fontface=if_else(in_package,"bold","italic"),
                     color=I(if_else(in_package,"#000000de","#ffffffde"))), 
                 family="Roboto Condensed") +
  coord_flip(ylim=c(0,6),xlim=c(-10,112)) 
  
  #ggsave(str_glue("2023/week-38/output/cran_color_sugiyama_{deg_cnt}.png"), device=ragg::agg_png, width=9*1.2, height=12*1.2)
  
  ggsave(str_glue("2023/week-38/output/cran_color_sugiyama_{deg_cnt}.svg"), device=svglite::svglite, width=9*1.2, height=12*1.2)
  
  
}


draw_by_grp(deg_cnt=0)

### draw by degree and produce separate images so I can combine them to make animation
possible_deg_cnt <- unname(possible_deg_cnt)
possible_deg_cnt[c(1:20)] |> rev() |> map(draw_by_grp)

### I feel like there's just so much efficient way to make animation.. but I can't figure out how just yet...



cran_color |> 
  left_join(cran_color_g |> as_tibble() |> select(package=name,deg) |>
              mutate(deg_color = colourvalues::color_values(deg,pal_mat))) |>
  select(deg,deg_color,package_url,package,published,title,description,depends,imports) |>
  mutate(package=str_glue("<h4><span style='background-color:{deg_color};color:{deg_color}'> ■ </span>  {package_url} 📦 </h4>")) |>
  arrange(deg) |>
  mutate(deg=if_else(is.na(deg),"Package with No Depenency - Not Included in Network Visualization",if_else(deg==1,"Package with 1 Dependency",str_c("Package with ",deg, " Dependencies")))) |>
  group_by(deg) |>
  gt() |>
  fmt_markdown(c(package)) |> 
  cols_hide(c(package_url, deg_color)) |>
  gt_theme_538() |> 
  tab_header(title="CRAN Package dealing with Colour Palettes",
             subtitle="Week 38 TidyTuesday Data Visualization Challenge")


```
